---
title: "HW2"
output: html_document
---



## Load the `GEOmetadb` library

```{r}
suppressMessages(library(GEOmetadb))
```


## Download the GEO database (if necessary)

```{r}
if(!file.exists("GEOmetadb.sqlite"))
{
  getSQLiteFile()
}
```


## Connect and list the tables

```{r}
geo_con <- dbConnect(SQLite(),'GEOmetadb.sqlite')
dbListTables(geo_con)
```


## List the fields available for gse and gpl

```{r}
dbListFields(geo_con, 'gse')
dbListFields(geo_con, 'gpl')
```


## Find all HCV gene expression data using the Illumina platform submitted by an investigator at Yale

```{r}
dbGetQuery(geo_con, 
    "SELECT 
        gse.title, 
        gse.gse, 
        gpl.gpl, 
        gpl.manufacturer, 
        gpl.description 
    FROM (gse JOIN gse_gpl ON gse.gse=gse_gpl.gse) j JOIN gpl ON j.gpl=gpl.gpl 
    WHERE gse.title like '%HCV%'
        AND gpl.manufacturer like '%Illumina%'
        AND gse.contact like '%Yale%';")
```


## Load the `data.table` library and convert all db tables to data.table tables (with just the information we need)

```{r}
suppressMessages(library(data.table))
gseDT <- data.table(dbGetQuery(geo_con, "SELECT gse, title FROM gse WHERE title like '%HCV%' AND contact like '%Yale%';"), key="gse")
gplDT <- data.table(dbGetQuery(geo_con, "SELECT gpl, description, manufacturer FROM gpl WHERE manufacturer like '%Illumina%';"), key="gpl")
gse_gplDT <- data.table(dbGetQuery(geo_con, "SELECT * from gse_gpl;"), key=c("gse", "gpl"))
```


## Reproduce the query using data tables

```{r}
merge(gplDT,gse_gplDT[gseDT],by="gpl")
```
